\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-10-27 doc/flexiblesusy-paper.tex
explain finding of analytic solutions to the EWSB equations
This addresses Reviewer 1, comment B3.

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  The electroweak symmetry breaking
conditions \eqref{eq:one-loop-ewsb-eq} are solved simultaneously using
the iterative multi-dimensional root finder algorithm
\code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific Library
(GSL).  If no root can be found, the
\code{gsl_multiroot_fsolver_hybrids} algorithm is tried as
alternative, which uses a variable step size but might be a little
slower.

>>>>

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by theses equations.  For
example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
determine $|\mu|$, while the sign of $\mu$ is not fixed.  Currently
\fs handles only real model parameters and renormalizable theories,
which restricts the number of possible multiple solutions of Eqs.\
\eqref{eq:one-loop-ewsb-eq} to be less or equal to four times the
number of equations.  In some cases, as for example in the real MSSM,
% where $\mu$ is chosen to be fixed by the EWSB equations
these different solutions might be related by one or more free global
signs in the form
%
\begin{align}
  \mu = \sign\mu \cdot \sqrt{\ldots} .
\end{align}
%
\fs recognizes such cases and introduces these global signs as
additional free parameters.  The case of the real MSSM, where only
$|\mu|$ is fixed by the EWSB equations, is therefore handled
automatically.  If, however, the different solutions are not related
by global signs, then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example done in the $Z_3$-violating NMSSM (\code{SMSSM}), where $\mu$
is chosen to be fixed by the EWSB equations: There the tree-level
solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with a non-global $\sign X$.  In this model the user can parametrize
the tree-level solution in terms of $\sign X$ in the
\code{TreeLevelEWSBSolution} variable and add $\sign X$ by hand to the
list of free input parameters, see
\code{model_files/SMSSM/FlexibleSUSY.m.in} for an example.  If the
user decides to not pick a particular solution by leaving the variable
\code{TreeLevelEWSBSolution} empty, \fs tries to find a solution to
the tree-level EWSB equations numerically via an iteration.  In this
case, however, the user does not have the option to chose between the
different solutions.

%
If a solution of the tree-level EWSB equations has been found, the
one-loop equations \eqref{eq:one-loop-ewsb-eq} are solved
simultaneously using the iterative multi-dimensional root finder
algorithm \code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific
Library (GSL).  If no root can be found, the
\code{gsl_multiroot_fsolver_hybrids} algorithm is tried as
alternative, which uses a variable step size but might be a little
slower.
